The polynomial representation of the type $A_{n - 1}$ rational Cherednik algebra in characteristic $p \mid n$

### 2015-05-18 formalwriteup.tex
\ell -> l

Then we see that for $0 \le \ell \le p$ we have
\[
[z^\ell](z^2V(z))=[z^\ell]G_0(z)=[z^\ell]G_1(z)=0.
\]
Then we also have
\[
[z^\ell](G_1(z)-G_0(z)+z^2V(z))=0.
\]
Then $[z^p]\left(\frac{G_1-G_0+z^2V}{1-x_0z}\right)$ is $0$ since it is a linear combination of $[z^\ell](G_1(z)-G_0(z)+z^2V(z))$ for $0 \le \ell \le p$, so 
\[
[z^p]\left((D_1-D_0)(F_0)\right)=[z^p]\left(-zF_0-z^2\frac{\partial F_0}{\partial z}\right).
\]
Let $b=[z^{p-1}](F_0)$. We see that $[z^p](-zF_0)=-b$. Then $[z^{p-2}]\left(\frac{\partial F_0}{\partial z}\right)=(p-1)b=-b$, so $[z^p]\left(-z^2\frac{\partial F_0}{\partial z}\right)=b$. Therefore $[z^p]\left(-zF_0-z^2\frac{\partial F_0}{\partial z}\right)=-b+b=0$, so 
\[
[z^p]((D_1-D_0)(F_0))=(D_1-D_0)([z^p]F_0)=(D_1-D_0)f_0=0.
\]
By the symmetry of the $F_i$, this is the only Dunkl operator we need consider; it is clear that all the $f_i$ are killed by the Dunkl operators.
\end{proof}

@@

\begin{lemma}\label{lem:vand}Let $y_0, \ldots, y_{r - 1}$ be distinct elements of $k$.  Then the polynomials $b_i(z) = \prod_{j \neq i} (1 - y_j z)$ are linearly independent.
\end{lemma}
\begin{proof} 
Suppose that for some $\lambda_i \in k$ we have $\sum_i \lambda_ib_i(z)=0$.  Then $\left(\prod_j (1-y_jz)\right)\left(\sum_i \frac{\lambda_i}{1-y_iz}\right)$ must be $0$ in the ring of formal power series. Since the ring of power series is an integral domain, $\sum_i \frac{\lambda_i}{1-y_iz}=0$. In particular for $\ell=0,\dots,r-1$ we have $\sum_i \lambda_iy_i^\ell=0$.  However, the vectors $(1,y_i,\dots,y_i^{r-1})$ are linearly independent by the Vandermonde determinant, so $\lambda_i=0$ for all $i$, meaning that $b_i(z)$ are linearly independent.
\end{proof}

@@

We will assume that $\prod_i y_i \ne 0$ in order to derive a contradiction and show that one of the $y_i$ is $0$. Since $\sum_{i=0}^{n-1} x_i=0$ we also have $\sum_{i=0}^{r-1} m_iy_i=0$; this enables us to compute the derivative $F'(z)$, which is 
\[
F'(z)=V(z)+\left(\sum_j \frac{-cx_j}{1-x_jz}F(z)\right)=V(z)+\left(\sum_j \frac{-cm_jy_j}{1-y_jz}F(z)\right),
\]
where $[z^\ell]V(z)=0$ for $0 \le \ell \le p$.

We will use a `residues by parts' argument; using the fact that the residue of an exact differential is zero, we will relate the coefficients of a power series to the coefficients of the derivative of that power series and derive a contradiction. 
Let
\[
a(z)=\frac{1}{\prod_{i=0}^{r-1}(1-y_iz)}F(z).
\]
We then see that
\begin{align*}
a'(z)&=\frac{d}{dz}\left(\frac{1}{\prod_{i=0}^{r-1}(1-y_iz)}F(z)\right)\\
&=\frac{1}{\prod_{i=0}^{r-1}(1-y_iz)}F'(z)+\sum_{j=0}^{r-1}\frac{y_j}{(1-y_jz)\prod_{i=0}^{r-1}(1-y_iz)}F(z)\\
&=\frac{1}{\prod_{i=0}^{r-1}(1-y_iz)}\left(V(z)+\left(\sum_j \frac{-cm_jy_j}{1-y_jz}F(z)\right)\right)+\sum_{j=0}^{r-1}\frac{y_j}{1-y_jz}a(z)\\
&=\frac{V(z)}{\prod_{i=0}^{r-1}(1-y_iz)}+\left(\sum_j \frac{-cm_jy_j}{1-y_jz}a(z)\right)+\sum_{j=0}^{r-1}\frac{y_j}{1-y_jz}a(z)\\
&=\frac{V(z)}{\prod_{i=0}^{r-1}(1-y_iz)}-\left(\sum_j \frac{(m_jc-1)y_j}{1-y_jz}a(z)\right).
\end{align*}
Since $[z^\ell]V(z)=0$ for $0 \le \ell \le p$ we see that $\frac{V(z)}{\prod_{i=0}^{r-1}(1-y_iz)}=z^pb(z)$ for some power series $b(z)$. Therefore
\[
a'(z)=z^pb(z)-\sum_j \frac{(m_jc-1)y_j}{1-y_jz}a(z)
\]
for some power series $b(z)$.

@@

For any integer $\ell$, define the Laurent polynomial 
\[
h_\ell(z)=z^{-\ell-r}\prod_i(1-y_iz).
\]
Notice that
\begin{align*}
\frac{d}{dz}( h_\ell(z) a(z)) &= -(l + r)z^{-1} h_\ell(z) a(z) - \sum_i \frac{y_i}{1 - y_iz} h_\ell(z) a(z) + h_\ell(z) a'(z) \\
&=  -(\ell + r)z^{-1} h_\ell(z) a(z) - \sum_i \frac{y_i}{1 - y_iz} h_\ell(z) a(z) - h_\ell(z) \sum_j \frac{(m_j c - 1)y_i}{1 - y_ijz} a(z)  + z^ph_\ell(z) b(z)\\
&= -(\ell + r)z^{-1} h_\ell(z) a(z) - \sum_i \frac{y_im_i c}{1 - y_iz} h_\ell(z) a(z) + z^ph_\ell(z) b(z)\\
&= -\left((\ell + r) z^{-1} + \sum_i \frac{y_im_i c}{1 - y_iz}\right) h_\ell(z) a(z) + z^ph_\ell(z) b(z).
\end{align*}
We note that 
\[
-\left((\ell + r) z^{-1} + \sum_i \frac{y_im_i c}{1 - y_iz}\right) h_\ell(z)
\]
is a Laurent polynomial with lowest degree term 
\[
- z^{-l-r-1} \left(\ell + r + \sum_i y_i m_i c\right) = - z^{-\ell-r-1}(\ell+ r)
\]
and highest degree term 
\[
- z^{-\ell - 1} (-1)^r  \prod_i y_i\left((\ell + r) - \sum_i m_i c\right).
\]
The lowest degree term of the formal power series $z^ph_\ell(z)b(z)$ has degree at least $p-\ell-r$. We note that we must have $[z^{-1}]\left(\frac{d}{dz}( h_\ell(z) a(z))\right)=0$, so for $p-\ell-r \ge 0$, we see that $[z^{-1}]\left(\frac{d}{dz}( h_\ell(z) a(z))\right)=0$ is a linear combination of $[z^\ell]a(z),[z^{\ell+1}]a(z),\dots,[z^{\ell+r}]a(z)$ in which $[z^\ell]a(z)$ has nonzero coefficient $ (-1)^r  \prod_i y_i\left((\ell + r) - \sum_i m_i c\right)$. The latter coefficient is nonzero since $c$ is generic and none of the $y_i$ are $0$. Therefore $[z^\ell]a(z)$ is a linear combination of $[z^{\ell+1}]a(z),\dots,[z^{\ell+r}]a(z)$ when $\ell \le p-r$.

Since $[z^{p-r+1}]a(z)=\dots=[z^p]a(z)=0$, by induction starting with $\ell=p-r$ we find that $[z^\ell]a(z)=0$ for $\ell=p-r,p-r-1,\dots,0$. Therefore, $[z^0]a(z)=0$; a contradiction. We conclude that $y_i=0$ for some $i$.

>>>>

Then we see that for $0 \le l \le p$ we have
\[
[z^l](z^2V(z))=[z^l]G_0(z)=[z^l]G_1(z)=0.
\]
Then we also have
\[
[z^l](G_1(z)-G_0(z)+z^2V(z))=0.
\]
Then $[z^p]\left(\frac{G_1-G_0+z^2V}{1-x_0z}\right)$ is $0$ since it is a linear combination of $[z^l](G_1(z)-G_0(z)+z^2V(z))$ for $0 \le l \le p$, so 
\[
[z^p]\left((D_1-D_0)(F_0)\right)=[z^p]\left(-zF_0-z^2\frac{\partial F_0}{\partial z}\right).
\]
Let $b=[z^{p-1}](F_0)$. We see that $[z^p](-zF_0)=-b$. Then $[z^{p-2}]\left(\frac{\partial F_0}{\partial z}\right)=(p-1)b=-b$, so $[z^p]\left(-z^2\frac{\partial F_0}{\partial z}\right)=b$. Therefore $[z^p]\left(-zF_0-z^2\frac{\partial F_0}{\partial z}\right)=-b+b=0$, so 
\[
[z^p]((D_1-D_0)(F_0))=(D_1-D_0)([z^p]F_0)=(D_1-D_0)f_0=0.
\]
By the symmetry of the $F_i$, this is the only Dunkl operator we need consider; it is clear that all the $f_i$ are killed by the Dunkl operators.
\end{proof}

@@

\begin{lemma}\label{lem:vand}Let $y_0, \ldots, y_{r - 1}$ be distinct elements of $k$.  Then the polynomials $b_i(z) = \prod_{j \neq i} (1 - y_j z)$ are linearly independent.
\end{lemma}
\begin{proof} 
Suppose that for some $\lambda_i \in k$ we have $\sum_i \lambda_ib_i(z)=0$.  Then $\left(\prod_j (1-y_jz)\right)\left(\sum_i \frac{\lambda_i}{1-y_iz}\right)$ must be $0$ in the ring of formal power series. Since the ring of power series is an integral domain, $\sum_i \frac{\lambda_i}{1-y_iz}=0$. In particular for $l=0,\dots,r-1$ we have $\sum_i \lambda_iy_i^l=0$.  However, the vectors $(1,y_i,\dots,y_i^{r-1})$ are linearly independent by the Vandermonde determinant, so $\lambda_i=0$ for all $i$, meaning that $b_i(z)$ are linearly independent.
\end{proof}

@@

We will assume that $\prod_i y_i \ne 0$ in order to derive a contradiction and show that one of the $y_i$ is $0$. Since $\sum_{i=0}^{n-1} x_i=0$ we also have $\sum_{i=0}^{r-1} m_iy_i=0$; this enables us to compute the derivative $F'(z)$, which is 
\[
F'(z)=V(z)+\left(\sum_j \frac{-cx_j}{1-x_jz}F(z)\right)=V(z)+\left(\sum_j \frac{-cm_jy_j}{1-y_jz}F(z)\right),
\]
where $[z^l]V(z)=0$ for $0 \le l \le p$.

We will use a `residues by parts' argument; using the fact that the residue of an exact differential is zero, we will relate the coefficients of a power series to the coefficients of the derivative of that power series and derive a contradiction. 
Let
\[
a(z)=\frac{1}{\prod_{i=0}^{r-1}(1-y_iz)}F(z).
\]
We then see that
\begin{align*}
a'(z)&=\frac{d}{dz}\left(\frac{1}{\prod_{i=0}^{r-1}(1-y_iz)}F(z)\right)\\
&=\frac{1}{\prod_{i=0}^{r-1}(1-y_iz)}F'(z)+\sum_{j=0}^{r-1}\frac{y_j}{(1-y_jz)\prod_{i=0}^{r-1}(1-y_iz)}F(z)\\
&=\frac{1}{\prod_{i=0}^{r-1}(1-y_iz)}\left(V(z)+\left(\sum_j \frac{-cm_jy_j}{1-y_jz}F(z)\right)\right)+\sum_{j=0}^{r-1}\frac{y_j}{1-y_jz}a(z)\\
&=\frac{V(z)}{\prod_{i=0}^{r-1}(1-y_iz)}+\left(\sum_j \frac{-cm_jy_j}{1-y_jz}a(z)\right)+\sum_{j=0}^{r-1}\frac{y_j}{1-y_jz}a(z)\\
&=\frac{V(z)}{\prod_{i=0}^{r-1}(1-y_iz)}-\left(\sum_j \frac{(m_jc-1)y_j}{1-y_jz}a(z)\right).
\end{align*}
Since $[z^l]V(z)=0$ for $0 \le l \le p$ we see that $\frac{V(z)}{\prod_{i=0}^{r-1}(1-y_iz)}=z^pb(z)$ for some power series $b(z)$. Therefore
\[
a'(z)=z^pb(z)-\sum_j \frac{(m_jc-1)y_j}{1-y_jz}a(z)
\]
for some power series $b(z)$.

@@

For any integer $l$, define the Laurent polynomial 
\[
h_l(z)=z^{-l-r}\prod_i(1-y_iz).
\]
Notice that
\begin{align*}
\frac{d}{dz}( h_l(z) a(z)) &= -(l + r)z^{-1} h_l(z) a(z) - \sum_i \frac{y_i}{1 - y_iz} h_l(z) a(z) + h_l(z) a'(z) \\
&=  -(l + r)z^{-1} h_l(z) a(z) - \sum_i \frac{y_i}{1 - y_iz} h_l(z) a(z) - h_l(z) \sum_j \frac{(m_j c - 1)y_i}{1 - y_ijz} a(z)  + z^ph_l(z) b(z)\\
&= -(l + r)z^{-1} h_l(z) a(z) - \sum_i \frac{y_im_i c}{1 - y_iz} h_l(z) a(z) + z^ph_l(z) b(z)\\
&= -\left((l + r) z^{-1} + \sum_i \frac{y_im_i c}{1 - y_iz}\right) h_l(z) a(z) + z^ph_l(z) b(z).
\end{align*}
We note that 
\[
-\left((l + r) z^{-1} + \sum_i \frac{y_im_i c}{1 - y_iz}\right) h_l(z)
\]
is a Laurent polynomial with lowest degree term 
\[
- z^{-l-r-1} \left(l + r + \sum_i y_i m_i c\right) = - z^{-l-r-1}(l+ r)
\]
and highest degree term 
\[
- z^{-l - 1} (-1)^r  \prod_i y_i\left((l + r) - \sum_i m_i c\right).
\]
The lowest degree term of the formal power series $z^ph_l(z)b(z)$ has degree at least $p-l-r$. We note that we must have $[z^{-1}]\left(\frac{d}{dz}( h_l(z) a(z))\right)=0$, so for $p-l-r \ge 0$, we see that $[z^{-1}]\left(\frac{d}{dz}( h_l(z) a(z))\right)=0$ is a linear combination of $[z^l]a(z),[z^{l+1}]a(z),\dots,[z^{l+r}]a(z)$ in which $[z^l]a(z)$ has nonzero coefficient $ (-1)^r  \prod_i y_i\left((l + r) - \sum_i m_i c\right)$. The latter coefficient is nonzero since $c$ is generic and none of the $y_i$ are $0$. Therefore $[z^l]a(z)$ is a linear combination of $[z^{l+1}]a(z),\dots,[z^{l+r}]a(z)$ when $l \le p-r$.

Since $[z^{p-r+1}]a(z)=\dots=[z^p]a(z)=0$, by induction starting with $l=p-r$ we find that $[z^l]a(z)=0$ for $l=p-r,p-r-1,\dots,0$. Therefore, $[z^0]a(z)=0$; a contradiction. We conclude that $y_i=0$ for some $i$.
